### 2016-03-21 ch03-lrm/LPMSmooth_ML_EG_JS_JL.tex
Some figures reimported as TikZ + text on truncation

\researchBasedOn[This section]{Lumori2014TIM}
\TODO{figures to TikZ}
\TODO{re-introduce $F$-test truncator and box-carr averager}
\TODO{labels into standardized form}

@@

\begin{figure}
\centering
\includegraphics{\thisDir/figs/tikz0.pdf}
\caption{SISO LTI discrete-time system in an output error setup.}
\label{lpmtdrep}
\end{figure}

@@

\item
Assuming that the impulse response decays exponentially in time, the noise is bound to predominate after a certain time interval.

@@

This action results in smoothing the \gls{LPM} estimate of the \gls{FRF}.
An exponential fitting method is introduced to determine the truncation point.
\end{enumerate}

@@

\begin{align}
\hat g_\mathrm{poly}(t) = \hat g_{\mathrm{poly}\setminus \mathrm{DC}}(t) - m_{g10},\ \text{for}\ t=0,1,\dots,N-1
\end{align}

\subsection{Truncation of the Impulse Response Via Exponential Fitting}\label{se:truncImpulseResp}

@@

\begin{figure}[tbh] %top bottom here
    \centering
    \includegraphics[scale=0.9]{\thisDir/figs/figLPMvsTruncWlegend}
    \caption[Comparison of FRF estimated using LPM and Truncated LPM.]{Comparison of the \gls{LPM} and Truncated \gls{LPM} estimates of the \gls{FRF} for an \gls{SNR} of $18.3\unit{dB}$. The gray plots pertain to the \gls{LPM} estimate without truncation, the black ones to the \gls{LPM} with truncation (trunc). Dots: \gls{FRF} estimates $\hat{G}$. Crosses: sample standard deviations $\hat{\sigma}$. Dashed lines: bias $\hat{b}$. White line: the true system $G_0$.}
    \label{figLPMvsTrunc}
\end{figure}

@@

\begin{figure}[htb] %  figure placement: here, top, bottom, or page
   \centering
	\includegraphics{\thisDir/figs/HistogramExpFitting.pdf}
         \caption[RMS error of the FRF versus truncation time.]{Black full line (left hand side y-axis (ordinate)): total \gls{RMS} error as a function of the truncation time (the single black dot is the minimum of the \gls{RMS}). Gray histogram (right hand side y-axis (ordinate)): selected truncation times for 1000 noise realizations by the exponential fitting methods.}
   \label{fig:pdfAndRMSeVStruncTime}
\end{figure}

@@

The truncation time is determined from the data as described in \secref{se:truncImpulseResp}. This was done on $1000$ realizations of the noise, and depicted in \figref{fig:pdfAndRMSeVStruncTime} by the histogram.
Clearly, the method for selecting the truncation time $t_{\mathrm{trunc}}$ has a good overall performance, based on the mode of its distribution. A closely grouped set of values for $t_{\mathrm{trunc}}$ is obtained around the $90^{\text{th}}$ sample, without significant outliers.

From the plot, we can conclude that the \gls{RMSE} increases rapidly when $t_{\mathrm{trunc}}$ is smaller than the optimal.
On the other hand, selection of a value of $t_{\mathrm{trunc}}$ that is too large, is not nearly as detrimental to the modeling error.
The graph also shows that the \gls{RMSE} of the model can be decreased from $0.62$ (without truncation) to $0.18$ when the optimal truncation is applied. Also, one observes a low sensitivity of the \gls{RMSE} w.r.t.~$t_\mathrm{trunc}$, when truncating at times beyond that optimum. Therefore, a somewhat conservative truncation method is still likely to yield a close to optimal result.

\subsection{Conclusion}\label{se:conclusion}

>>>>

\researchBasedOn[This section]{Lumori2014TIM}
\TODO{figures to TikZ}
\TODO{labels into standardized form}

@@

\begin{figure}
\centering
\setlength\figurewidth{0.85\columnwidth}
\setlength\figureheight{0.68\figurewidth}
\input{\thisDir/figs/oesetup.tikz}
\caption{SISO LTI discrete-time system in an output error setup.}
\label{lpmtdrep}
\end{figure}

@@

\item Assuming that the impulse response decays with the passing of time, the noise is bound to predominate after a certain time interval. 
In particular, when $t > \truncTime$, the impulse response has decayed below the noise floor.
In \secref{sec:nonparametric:truncation:Ftest} and \secref{sec:nonparametric:truncation:exponentialFit}, two approaches are introduced to estimate $\truncTime$.

@@

This action results in smoothing the \gls{LPM} estimate of the \gls{FRF}.
\end{enumerate}

@@

\begin{align}
\hat g_\mathrm{poly}(t) = \hat g_{\mathrm{poly}\setminus \mathrm{DC}}(t) - m_{g10},\ \text{for}\ t=0,1,\dots,N-1
\end{align}
\end{enumerate}

\subsection{Truncation Via the $F$-Test}
\label{sec:nonparametric:truncation:ftest}
To distinguish the part of the measured impulse response where the actual impulse response dominates from the part where the noise dominates, the estimated impulse response $\hat{g}_\mathrm{poly}$ is split into smaller segments. 
One can then gradually determine whether these segments are noise or the actual impulse response.

% 1) cut into NS segments of minimally NM samples
% 2) assume last segment == noise
%    start from back
% 3) F-test on Var, Var==RMS if dc == 0 (i.e. noise)
%    H0: s²(1) =  s²(2)
%    H1: s²(1) >  s²(2)
%    alpha >> ==> beta <<
% 4) if H1 and RMS(signal)/RMS(noise) > f
%    robustness, SNR
%    select segment i
% 5) refine until wanted resolution is obtained

\begin{enumerate}
  \item The impulse response $\hat{g}_\mathrm{poly}(t)$ is split into $N_S$ segments of approximate equal lengths
$L_S = \floor{\frac{N}{N_S}}$. 
If necessary, the length of the first segment is reduced to allow for the others to have equal lengths.
The $i^{\text{th}}$ segment is denoted $\segm{\hat{g}}{i}$, for $i \in \left\{1,\ldots,N_S\right\}$.
  By \assref{ass:decay90perctime}, the last segment, $\segm{\hat{g}}{N_S}$, is noise when $L_S \leqslant 0.1N$.

  \item Iteration is then carried out from the last segment ($i = N_S$) towards the beginning of the impulse response. To determine whether the preceding segment $\segm{\hat{g}}{i-1}$ is completely dominated by noise,  both an $F$-test and an absolute criterion on the RMS are used.

  The $F$-test is suited to compare the variances of different segments $\segm{\hat{g}}{i}$ and $\segm{\hat{g}}{i-1}$ \citep{Parsons1974}, while accounting for the degrees of freedom in each segment.
  For zero-mean noise, these variances are the \gls{RMS} value of each noise segment.
  This has the advantage that a segment in which the impulse response still has a significant amplitude is likely to have a large \gls{RMS} value, which is more likely to be detected.
  Let $s^2(i)$ denote the mean square value of segment $\segm{\hat{g}}{i}$, \emph{viz}:
  \begin{equation}
    s^2(i) 
    \isdef \frac{1}{L_S} \sum_{j=1}^{L_S} \left(\segm{\hat{g}}{i}(j)\right)^2
    = \left( \rms{\hat{g}^{[i]}} \right)^2
           \text{.}
  \end{equation}
  Starting with the last segment ($i = N_S$), the following null hypothesis and the alternative hypothesis are formulated:
  \begin{align}
     \nullHypothesis &: s^2(i-1) = s^2(i)\\
     \altHypothesis    &: s^2(i-1) > s^2(i)
     \text{.}
  \end{align}
  As segment $\segm{\hat{g}}{i}$ has been tested and been classified as noise, the null hypothesis states that the segment $\segm{\hat{g}}{i-1}$ is also noise.
  The alternative case occurs when $\segm{\hat{g}}{i-1}$ has a significant change, which indicates the presence of the signal.
  Using the $F$-test on the \gls{RMS} values of the two segments with a high confidence level of $\alpha=0.999$ and the aforementioned hypotheses, we can determine whether $\segm{\hat{g}}{i-1}$ is likely to be part of the actual impulse response.
  In particular, the null hypothesis is rejected when the empirical $F$ statistic is small, i.e.
  \begin{equation}
    F \isdef \frac{s^2(i-1)}{s^2(i)} \leq \ICDF{\FDistribution} \left( \alpha; n_{i-1}, n_{i} \right)
  \end{equation}
  where $ \ICDF{\FDistribution}$ is the inverse \gls{CDF} of an $F$ distribution with $n_{i-1}$ and $n_{i}$ degrees of freedom.
  In this case $n_{i-1}$ and $n_{i}$ are the respective number of samples in the $(i-1)^{\text{th}}$ and $i^{\text{th}}$ segment of the impulse response.

  The high level of confidence ($\alpha=0.999$) ensures that the probability of a Type~II error is smaller than $1-\alpha$ \citep{Parsons1974}.
  In our case, such an error means that a part of the actual impulse response would be falsely classified as noise, which could significantly increase the bias as information of the system is discarded.
  A Type~I error is less detrimental since, in that case, noise is falsely classified as a signal component and kept in the impulse response, thereby causing a sub-optimal value of the variance of the estimate.
  As the \gls{LPM} samples are correlated over a short frequency span, the actual noise present in $\hat{g}$ may be slightly non-stationary.
  To cope with this, one can introduce other criteria which must be satisfied together with the outcome of the $F$-test.
  A criterion that shows good results is to check whether the segment $\segm{\hat{g}}{i-1}$ has an RMS value that is at least a factor $\kappa$ larger than the RMS of the noise.
  Even for a moderate $\kappa = 1.1$, a large improvement in the detection was observed. 
%  \JL{Q: doesn't this increase the probability of a Type II error?}
%  \EG{A: This slightly increases for small $\kappa$ and increases significantly for large values of $\kappa$. But it is needed to overcome nonstationarity of the noise.}

  \item This procedure is repeated until a segment $\segm{\hat{g}}{i-1}$ is found that is dominated by the actual impulse response according to the outcome of the $F$-test and the absolute \gls{RMS} value of the segment.
  At that point, it is very likely that the transition from noise to the actual impulse response happens within the segment $\segm{\hat{g}}{i-1}$.
  One now has the choice to accept the last sample of $\segm{\hat{g}}{i-1}$ to be the last meaningful sample $\truncTime$ for the impulse response.
  The accuracy of this estimate is limited by the length of the segment $L_S$.

  \item
  A more accurate estimate can be obtained by dividing the segment $\segm{\hat{g}}{i-1}$ -- which may contain both the signal and noise -- yet again into smaller segments. The procedure described above is then repeated until a satisfactory, accurate $\truncTime$ is obtained, or until the subsegments have a length that is too short ($L_{S\min} = 10$) to guarantee the accuracy of the RMS values. %rely on the RMS values to be accurate.
  To start this refinement, the last $\segm{\hat{g}}{i}$ should be used to compare the RMS with the $F$-test, since the last subsegment of $\segm{\hat{g}}{i-1}$ cannot be asserted to be dominated by noise.

\end{enumerate}
This procedure is illustrated in \figref{fig:nonparametric:truncation:Ftest} for the system described by the following simulation equations, which are relevant to the LPM outlined in Section \ref{se:LPMFRFest}:
\begin{subequations}
\label{eq:systemSimulations}
\begin{align}
y_0(t)  &= 1.5371y_0(t-1)    -0.9025y_0(t-2) + u(t)\\
y(t)       &= y_0(t) + e(t),
\end{align}
\end{subequations}
where $e(t)$ is a white noise sequence, such that the SNR of the output signal is 78.2~dB.
% \eqref{eq:systemSimulations}. 
The figure shows segments of the estimated impulse response. The last two segments are dominated by the noise, while all samples at $t<80$ have large system contributions. The algorithm outlined above selects $\truncTime = 80$, beyond which all samples are set to zero, resulting in a smooth estimate.  %after which all samples are fixed to a zero value for the smoothed estimate.

\begin{figure}[tbh]
\centering
\remakeNextFigure
 \setlength\figurewidth{0.8\columnwidth}
 \setlength\figureheight{0.68\figurewidth}
  \input{\thisDir/figs/truncRMS.tikz}
\caption[Illustration of impulse response truncation using the $F$-test.]{Illustration of truncating the impulse response by performing an $F$-test on successive segments of the impulse response. The different segments are separated by dotted lines.}
\label{fig:nonparametric:truncation:Ftest}
\end{figure}

\subsection{Truncation of the Impulse Response Via Exponential Fitting}
\label{sec:nonparametric:truncation:exponentialfit}

@@

\begin{figure}[tbh] %top bottom here
    \centering
    \setlength\figurewidth{0.85\columnwidth}
    \setlength\figureheight{0.68\figurewidth}

    \includegraphics[scale=0.9]{\thisDir/figs/figLPMvsTruncWlegend}
    \caption[Comparison of FRF estimated using LPM and Truncated LPM.]{Comparison of the \gls{LPM} and Truncated \gls{LPM} estimates of the \gls{FRF} for an \gls{SNR} of $18.3\unit{dB}$. The gray plots pertain to the \gls{LPM} estimate without truncation, the black ones to the \gls{LPM} with truncation (trunc). Dots: \gls{FRF} estimates $\hat{G}$. Crosses: sample standard deviations $\hat{\sigma}$. Dashed lines: bias $\hat{b}$. White line: the true system $G_0$.}
    \label{figLPMvsTrunc}
\end{figure}

@@

\begin{figure}[htb] %  figure placement: here, top, bottom, or page
   \centering
        \setlength\figurewidth{0.85\columnwidth}
        \setlength\figureheight{0.68\figurewidth}
        \input{\thisDir/figs/pdfs.tikz}
         \caption[RMS error of the FRF versus truncation time.]{
         The left y-axis shows the \gls{RMS} error of the estimated \gls{FRF} as a function of the chosen truncation time $\truncTime$.
         The corresponding optimal $\truncTime$ is indicated as well.
         Both bar plots (right y-axis) indicate for $\nMC =1000$ realizations of the disturbing noise and excitation signal, the number of times that $\truncTime$ has been selected by means of either the method using the $F$-test or the method that fits an exponential.
         It can be seen that both methods allow to almost halve the \gls{RMS} error and approach the optimal truncation time.}
   \label{fig:pdfAndRMSeVStruncTime}
\end{figure}

@@

The truncation time is determined from the data as described in \secref{sec:nonparametric:truncation:exponentialfit}. This was done on $1000$ realizations of the noise, and depicted in \figref{fig:pdfAndRMSeVStruncTime} by the histogram.
Clearly, both methoda for selecting the truncation time $\truncTime$ have a good overall performance, based on the mode of its distribution (around the $90^{\text{th}}$ sample).

For the $F$-test based method, the obtained $\truncTime$ values only take a discrete set of values due to the limitation of the segment length.
While most trials result in $\truncTime \in \Set{90, 80, 70}$, a few outliers can be observed.
This can be attributed to the high power of the $F$-tests that are performed.

For the exponential fitting approach, however, a closely grouped set of (continuous) values for $\truncTime$ is obtained around the $90^{\text{th}}$ sample, without significant outliers.
Although this distribution has a mode quite a bit higher than the method using the $F$-test, due to the absence of outliers, this method is more reliable to use.

From the plot, we can conclude that the \gls{RMSE} increases rapidly when $\truncTime$ is smaller than the optimal.
On the other hand, selection of a value of $\truncTime$ that is too large, is not nearly as detrimental to the modeling error.
The graph also shows that the \gls{RMSE} of the model can be decreased from $0.62$ (without truncation) to $0.18$ when the optimal truncation is applied. Also, one observes a low sensitivity of the \gls{RMSE} w.r.t.~$t_\mathrm{trunc}$, when truncating at times beyond that optimum. 
Therefore, a somewhat conservative truncation method is still likely to yield a close to optimal result.

\begin{guideline}[Check and reinforce assumptions in the data.]
As illustrated by the presented smoothing operation, it pays off to check typical assumptions of the system (e.g. smoothness of the \gls{FRF}, finite support of the impulse response) and even artificially reinforce such assumptions to improve the model quality.
\end{guideline}

\subsection{Conclusion}
\label{se:conclusion}
